Realistically Interesting Domains for Probabilistic Planners

### 2014-04-13 FinalPaper/paper.tex
adjusted on comments

\begin{abstract}

\todo{Rewrite abstract - Imara}

	In contrast to current demand for probabilistically interesting domains in the International Probabilistic Planning Competitions, this papier argues for using real-world appropriate domains. By relaxing the definition of a dead-end, we argue that for many
real life domains a simpler planning algorithm can be sufficient and even better than sophisticated probabilistic planners.

In this paper we evaluate the strength of FF-Replan, winner of IPPC-04 and unenrolled winner of IPPC-06, a replanner that uses the classical planner FF on a determinized probabilistic problem against a state of the art probabilistic planner \textsc{Gourmand}, the successor of the Labelled Real-Time Dynamic Probabilistic planner \textsc{Glutton}, runner up of the IPPC-11. By translating the Navigation and Traffic domain from IPPC-11 to a matching real-world equivalent and evaluating these using the proposed probabilistically interesting baseline test as proposed by Little et al\. we evaluate whether FF-Replan is in fact capable enough to solve enough of these so called classical planning problems in disguise.
\end{abstract}

@@

\todo{iterate on background - Imara}
\subsection{Planning problems}

Planning is the process of computing several steps of a problem-solving procedure before executing any of them.

\subsubsection{Classical Planning problems.}

A classical planning problem is defined as states, actions, conditions and a goal state and are solved using search and can be aided using heuristics. The environment is assumed to be \emph{deterministic, observable and static}. The planner tries to find a sequence of actions such that a goal or maximum reward is achieved.

\subsubsection{Probabilistic Planning problems}

Real-World applications are far from deterministic, induced by natural dynamics and chance. To accommodate for Real-World domains probabilistic planning problems were introduced.  These systems can be described using \emph{Markov Decision Processes} (MDPs) and can account for observable uncertainty. Probabilistic planning problems can be solved by evaluating iterations and creating policies to reach a goal state, either online, while executing the plan, or offline, where all possibilities is calculated upfront.

\subsection{Problem domains}

\subsubsection{Goal-oriented MDPs. }The problems in IPPC-04 were specified using the Planning Domain Description Language (PPDLL). PPDDL specifies a probabistic planning problem as a set of probabilistic action sets and an initial and goal state description in which every action has a certain probability assigned to it \cite{FFReplan}. These MDPs are of the form $\mathit{M = (S, A, app, Pr, R)}$ where $\mathit{S}$ and $\mathit{A}$ are a finite set of states and actions, $\mathit{app(s)}$ is the set of all actions applicable in $\mathit{s}$. $\mathit{Pr}$ represents the probability of a state transaction $\mathit{(s, a, s')}$ for $\mathit{a}$ and $\mathit{R(s, a, s')}$ is the reward of applying $\mathit{a}$ to $\mathit{s}$ and generating $\mathit{s'}$ The planning problem is defined as a triple $\mathit{P = (M, s\textsubscript{0}, G, \rho)}$ where $\mathit{M = (S, A, app, Pr, R)}$ is an MDP, $\mathit{s\textsubscript{0} \in S}$ is the initial state, $\mathit{G \subseteq S}$ is the set of goal states and $\mathit{0 < \rho < 1}$ is a probability threshold \cite{teichteil2010incremental}.

 The planners from IPPC-04 were evaluated on how many trials reached the goal per problem and the average number of actions selected when successful. These planning problems were however relatively easy to solve as they did not account for \emph{natural dynamics} and/or large branching factors that occur in more realistic systems \cite{kolobov2012glutton}. FF-Replan \cite{FFReplan}, as specified in Section \ref{sec:planners} kindly remembered the committee of the IPPC-04 that the specified planning problems that contained no dead-ends were fairly easy to solve using relatively simple tools to handle with uncertainty.

\subsubsection{MDPs with Finite Horizon. }IPPC-11 decided to respond by introducing more realistic MDPs and their evaluation criteria by using MDPs with finite horizons. The planning problems are defined as tuples of the form \\ $\mathit{M\textsubscript{s0}(H) = \langle \langle S, A, T, R, s\textsubscript{0} \rangle , H \rangle }$ where $\mathit{S}$ s a finite set of states, $\mathit{A}$ a finite set of actions, $\mathit{T}$ is the transition function, similar to $\mathit{Pr}$ in the goal oriented MDP. $\mathit{R}$ represents the transitioning reward, $\mathit{s\textsubscript{0}}$ and $\mathit{H}$ is the number of decision epochs after which the process stops. Solving $\mathit{M(H)}$ means finding a policy for executing recommended state from the initial state over $\mathit{H}$ decision epochs \cite{kolobov2012gourmand}.

Translating finite horizon MDPs to goal-oriented domains results ends up in all state $H$ steps away from $s\textsubscript{0}$ are goal states. The evaluation criteria in finite horizon MDPs is the \emph{expected reward} and not the probability of the policy reaching the goal states.

\subsection{Planning algorithms}
\subsubsection{Replanning}
A replanner is a planner that takes a problem as input, plans a state-action policy, executes it online and whenever it encounters a state that was not in its state-action policy, it takes the current state as starting state and plans again until it has reached a goal or dead-end state. For our paper have chosen to evaluate FF-Replan \cite{FFReplan}, the winner of IPPC-04 and unenrolled winner of IPPC-06 as an inspiring, fairly simple planner that in our opinion is capable of employment in real-world domains.

\subsubsection{Probabilistic planning}
Probabilistic planners are used to solve probabilistic planning problems and should be able to account for uncertainty as opposed to classical planners. For our evaluation we have chosen to compare FF-Replan to a state-of-the-art probabilistic planner: \textsc{Gourmand}. \textsc{Gourmand} is the online successor of the \textsc{Glutton} plan, a Labelled Real-Time Dynamic Programming (LRTDP) based planner and the runner up of IPPC-11.

\subsection{Probabilistic Interestingness}
After IPPC-06, a baseline test for \emph{probabilistically interesting} planning domains was proposed by Little and Thi\'ebauw \cite{little2007probvsreplan}.  Problems that fail this test were deemed to be classical planning problems in disguise, favouring the use of replanners. A probabilistic planning problem is considered to be probabilistically interesting if and only if it has all the following structural properties \cite{little2007probvsreplan}:
% reference little & thiebauw again. Maybe also shortly define what all o, n, act etc. are.

\begin{itemize}
	\item there are multiple goal trajectories;
	\item there is at least on pair of distinct goal trajectories, $\tau$ and $\tau$ ', that share a common sequence of outcomes for the first n-1 outcomes, and where $\tau$\textsubscript{n} and $\tau$'\textsubscript{n} are distinct outcomes of the same action; and
	\item there exist two distinct goal trajectories $\tau$ and $\tau$' and outcome o $\in$ $\tau$ and o' $\in$ $\tau$' of two distinct actions a = act(o) and a'= act(o') such that executing a strictly decreases the maximum probability of reaching a state where a' can be executed.
\end{itemize}

@@

FF-Replan \cite{FFReplan}, a replanner based on determinization techniques and a classical planner, surprisingly outperformed other participating MDP solvers in the IPPC-04 and by doing so changed the course of the planning competition entirely. By relying on search and removing all probabilistic information from the problem, FF-replan functions with enormous speed and efficiency. This method has proven to outperform MDP solvers, the state-of-the-art probabilistic planners for uncertainty of that time. Since then, Little et al. has argued for \emph{probabilistically interesting} domains. By introducing finite horizon MDPs in IPPC-11, state of the art planners have been adapted to cope with large branching factors to approximate infinite planning problems. Because these ``planning problems'' can often be solved using a smaller horizon $h < H$ as introduced in Section \ref{sec:background}, they represent optimization problems rather than planning problems, favoring current state of the art planners over replanners. This section introduces the shift of the planners towards optimization  by discussing FF-replan and a current state of the art probabilistic planner \textsc{Gourmand} and its predecessors.

\subsection{FF-replan}
\label{sec:ffreplan}
 FF-Replan generates a deterministic planning problem by determinizing the input domain at the start of each planning iteration. This classical deterministic planning problem allows for the use of the classical FF planner \cite{Hoffmann01theff}. The FF planner then returns an ordered state-action plan for the deterministic problem of the current state. This plan is executed until an unexpected state is encountered in the following manner:

@@

FF-Replan has some obvious shortcomings. It removes all probabilistic information from the problem when determinizing the domain. The success of the planner is therefore fully reliable on the chosen determinization heuristic. As FF-replan uses no roll-out policy or future simulation, it has no way of avoiding \emph{avoidable} dead-ends. This explains why FF-replan performs poorly in domains containing dead-ends.

@@

FF-Replan replans online, an off-line planner like FF-Replan is RFF
\cite{teichteil2010incremental}. RFF generates policies by determinizing the
given MDP problem, builds partial policies off-line by producing partial
solution plans to the classical planning problem and incrementally aggregating
them into a policy, and using sequential Monte-Carlo simulations to asses the
probability of replanning. Experiments showed that RFF solved more problems
than FF-Replan because it takes the probability a policy has to replan into
account.

PROST \cite{keller2012prost} is a probabilistic planner, which contrary to
\textsc{Gourmand}, is UTC based. It uses techniques to reduce the branching
factor  and limits the search depth so it can focus more on parts that
influence the quality of the resulting policy. Reward lock detection is a
method to detect goals and dead ends in the finite horizon MDP\@. Q-value
Initialization improves the initial start of the UTC algorithm. In the executed
IPPC-11 benchmarks it outperforms other planners such as \textsc{Glutton}.
Later it has been surpassed by \textsc{Gourmand}.

\todo{rewrite prob int - Imara}

%\subsection{Probabilistic planning vs Replanning}
%Inspired by the FF-Replan entrance in the IPPC competition, this paper described a probabilistic interestingness quality of domains. It then compares probabilistic planners with replanners in the different domains to draw conclusions. Classical planning with replanning upon failure when there is so little or so much uncertainty about the world that it is not worth modelling and reasoning about or when goals dynamically change or when the problem is too large to solve at once. Probabilistic planners are more appropriate when avoiding irreparable or costly failures. Probabilistic planning computes a plan that can handle all foreseeable contingencies and is preferable for quick reactions in closed control loops.
%
%The following structural properties are identified for comparing probabilistic planning with replanning:
%\begin{itemize}
%	\item `dead' end states
%	\item the degree to which the probability of reaching a dead end state can be reduced through choice of actions
%	\item number of distinct trajectories
%	\item presence of mutual exclusion
%\end{itemize}
%
%If there are no dead ends, a replanner will always succeed. It is impossible to improve on the unavoidable dead ends (positive probability of reaching when executing which cannot be reduced without reducing the probability of reaching the goal).
%
%% repetition of what's already stated earlier
%
%A probabilistic planning problem is considered to be probabilistically interesting if and only if it has all the following structural properties:
%
%\begin{itemize}
%	\item there are multiple goal trajectories;
%	\item there is at least on pair of distinct goal trajectories, $\tau$ and $\tau$ ', that share a common sequence of outcomes for the first n-1 outcomes, and where $\tau$\textsubscript{n} and $\tau$'\textsubscript{n} are distinct outcomes of the same action; and
%	\item there exist two distinct goal trajectories $\tau$ and $\tau$' and outcome o $\in$ $\tau$ and o' $\in$ $\tau$' of two distinct actions a = act(o) and a'= act(o') such that executing a strictly decreases the maximum probability of reaching a state where a' can be executed.
%\end{itemize}
%
%Problems that fail to be \emph{probabilistically interesting} are actually classical planning problems in disguise and for those replanners are more appropriate. A replanner can also perform optimally for probabilistically interesting planning problems whenever the `most promising' goal trajectory is the correct thing to do. They create several domains to explore the issue of replanning vs\@. probabilistic planning.
%
%Following existing domains were said to be probabilistically interesting:
%\begin{itemize}
%	\item Drive
%	\item Exploding blocksworld
%	\item Pitchcatch
%	\item Schedule
%	\item Tireworld
%\end{itemize}
%
%They conclude by stating that the negative perception of probabilistic planning by the ICAPS community was inaccurate. They state that a synthesis of planning and replanning techniques could make a larger number of probabilistic planning problems practicably solvable than is currently possible.

>>>>

\begin{abstract}
By introducing and attempting to solve a planning problem domain involving Unmanned Aerial Vehicles (UAVs), this paper provides an overview of how the history of planning problem domains and their matching planners has changed throughout the scope of the International Planning Competitions (IPC). After the emergence of the FF-Replan planner in 2004, state-of-the-art benchmark domains have been argued to be more \emph{probabilistically interesting} causing a shift from the once dominant goal-oriented domains to the finite-horizon oriented domains.

This paper presents a currently hot topic domain in which several UAVs  work together to monitor a large area using restricted communication technology. To find an optimal planner for this domain, this paper presents an overview of other interesting domains and probabilistic planners that were involved in the redefinition of an interesting planning problem.

\end{abstract}

@@

\subsubsection{Planning Problems. } Planning is the process of computing several steps of a problem-solving procedure before executing any of them. A \emph{classical planning problem} is defined as states, actions, conditions and a goal state and is solved using when finding a state-action sequence from the initial state $s_0$ to a goal state $G$. To find a state-action sequence, the environment is assumed to be \emph{deterministic, observable and static}.

When planning for dynamic environments as is most often the case, natural dynamics and chance are introduced. To simulate these planning problems, the classical deterministic domains were extended by adding a \emph{probability} to every action, creating \emph{probabilistic planning problems}. These systems can be described using \emph{Markov Decision Processes} or MDPs.

\subsubsection{Goal-oriented MDPs. }Cclassical goal-oriented problems are specified using the Probabilistic Planning Domain Description Language (PPDDL). PPDDL specifies a probabilistic planning problem as a set of probabilistic actions and an initial and goal state description\cite{FFReplan}. These MDPs are of the form $\mathit{M = (S, A, app, Pr, R)}$ where $\mathit{S}$ and $\mathit{A}$ are a finite set of states and actions, $\mathit{app(s)}$ is the set of all actions applicable in the current state $\mathit{s}$. $\mathit{Pr}$ represents the probability of a state transition $\mathit{(s, a, s')}$ for action $\mathit{a}$ and $\mathit{R(s, a, s')}$ is the reward of applying $\mathit{a}$ to $\mathit{s}$ and generating $\mathit{s'}$ The planning problem is defined as a triple $\mathit{P = (M, s\textsubscript{0}, G}$ where $\mathit{M = (S, A, app, Pr, R)}$ is an MDP, $\mathit{s\textsubscript{0} \in S}$ is the initial state and $\mathit{G \subseteq S}$ is the set of goal states.

These planning problems were however relatively easy to solve as they did not account for large branching factors introduced by for instance decay \cite{kolobov2012glutton}. Natural dynamics cause the system to grow \myworries{exponentially??} as it introduces exogenous actions in every state. After FF-Replan \cite{FFReplan}, a replanner as introduced in Section  \ref{sec:planners} kindly reminded the committee of IPPC-04 that the specified planning problems containing no dead-ends were fairly easy to solve using relatively simple tools to handle uncertainty, the finite-horizon MDPs were introduces.

\subsubsection{MDPs with Finite Horizon. } Finite-horizon MDPs are an approximation for infinite planning problems using a finite horizon of a set amount of steps. Planning problems are defined as tuples of the form $\mathit{M(H) = \langle \langle S, A, T, R, s\textsubscript{0} \rangle , H \rangle }$ where $\mathit{S}$ is a finite set of states, $\mathit{A}$ a finite set of actions and $\mathit{T}$ is the transition function similar to $\mathit{Pr}$ in the goal oriented MDP. $\mathit{R}$ represents the transitioning reward, $\mathit{s\textsubscript{0}}$ the initial state and $\mathit{H}$ is the number of decision epochs after which the process stops. Solving $\mathit{M(H)}$ means finding a policy such that executing the recommended actions starting from state $(s_0, H)$ results in accumulating the largest expected reward over $H$ time steps \cite{kolobov2012glutton}.

\subsection{Planning algorithms}

These probabilistic MDPs can be solved using different kinds of planners. A \emph{replanner} is a planner that takes a problem as input, plans a partial state-action policy, executes it online and whenever it encounters a state that was not in its state-action policy, it takes the current state as starting state and plans again until it has reached a goal or dead-end state. To illustrate this technique, the notorious FF-planner is discussed in the next section.

\emph{Probabilistic planners} use sophisticated planning techniques to plan for uncertainty. By using simulations, roll-out policies or lookahead values, these planners prevent dead-ends and were therefore dominantly used for solving MDPs. To illustrate a probabilistic planner, the next section introduces \textsc{Gourmand}, an online state-of-the-art probabilistic planner, and its predecessors \textsc{Glutton} and the LRTDP based algorithms.

\subsection{Probabilistic Interestingness}
The \emph{probabilistically interesting} baseline test, proposed by Little and Thi\'ebauw \cite{little2007probvsreplan} after the success of FF-Replan initiated the shift of interest in goal-oriented MDPs to finite-horizon MDPs. Problems that fail this test are said to be classical planning problems in disguise, favoring the use of replanners. A probabilistic planning problem is considered to be probabilistically interesting if and only if it has all the following structural properties \cite{little2007probvsreplan}:

\begin{itemize}
	\item there are multiple goal trajectories;
	\item there is at least one pair of distinct goal trajectories, $\tau$ and $\tau'$, that share a common sequence of outcomes for the first $n-1$ outcomes, and where $\tau_n$ and $\tau'_n$ are distinct outcomes of the same action; and
	\item At least one of the distinct goal trajectories lead to outcomes o $\in$ $\tau$ and o' $\in$ $\tau$' of two distinct actions $a = act(o)$ and $a'= act(o')$ such that executing $a$ decreases the probability of reaching a dead end without influencing the probability of reaching a goal state. 
\end{itemize}

@@

FF-Replan \cite{FFReplan}, a replanner based on determinization techniques and a classical planner, surprisingly outperformed other participating MDP solvers in the IPPC-04 and by doing so changed the course of the planning competition entirely. By relying on search and removing all probabilistic information from the problem, FF-Replan functions with enormous speed and efficiency. This method has proven to outperform MDP solvers, the state-of-the-art probabilistic planners for uncertainty of that time. Since then, Little et al. has argued for \emph{probabilistically interesting} domains. By introducing finite horizon MDPs in IPPC-11, state of the art planners have been adapted to cope with large branching factors to approximate infinite planning problems. Because these ``planning problems'' can often be solved using a smaller horizon $h < H$ as introduced in Section \ref{sec:background}, they represent optimization problems rather than planning problems, favoring current state of the art planners over replanners. This section introduces the shift of the planners towards optimization  by discussing FF-Replan and a current state of the art probabilistic planner \textsc{Gourmand} and its predecessors.

\subsection{FF-ReFF-replanplan}
\label{sec:ffreplan}
 FF-Replan generates a deterministic planning problem by determinizing the input domain at the start of each planning iteration. This classical deterministic planning problem allows for the use of the classical FF planner \cite{Hoffmann01theff}. The FF planner then returns an ordered state-action plan for the deterministic problem of the current state. This plan is executed until an unexpected state is encountered in the following manner:

@@

FF-Replan has some obvious shortcomings. It removes all probabilistic information from the problem when determinizing the domain. The success of the planner is therefore fully reliable on the chosen determinization heuristic. As FF-Replan uses no roll-out policy or future simulation, it has no way of avoiding \emph{avoidable} dead-ends. This explains why  performs poorly in domains containing dead-ends.

@@

%In sections above we have described the GOURMAND planner and FF-Replan as two
%probabilistic planners of the IPPC competition.
%RFF~\cite{teichteil2010incremental} is an offline planner that generates a
%policy: for each reachable state it determines a plan to the goal. Like
%FF-Replan it uses determinization and FF\. FF returns an initial plan from the
%initial state to a goal state. For each state in the plan the successor states
%are added to the structure, after which Monte-Carlo sampling is used. States
%where the probability that they're reached is too low (below a threshold) are
%removed. The successor states are used as initial states in the next step. If
%FF fails to find a plan to the goal from this new state, the state is removed
%from the structure too. Results showed that RFF outperformed other competitors.
%A higher threshold means that RFF will generate more policies that are prone to
%replanning.

\subsubsection{PROST ?}
\todo{wrote sentence about Prost - Arian}\\
